\LARGE \bf {\sc underworlds}: Cascading Situation Assessment for Robots

### 2018-02-26 paper.tex
Update on Overleaf.
Large clean-up, save ~2 pages

However, {\sc Spark} is a monolithic module that does not support sharing its
internal 3D model with other external components. In contrast, \uwds focuses on
offering a shared and distributed representation of the environment within the
robot's software architecture. This also distinguishes \uwds from complex
cognitive toolkits like KnowRob (as found in OpenEASE~\cite{beetz2015open}).
While these tools maintain a spatio-temporal model of the world, this model is
internal and not meant to be made widely accessible to other external processes.
\uwds focuses instead on reusability and sharing of distributed spatio-temporal
models.  As such, \uwds can be seen as a middleware for spatio-temporal world
models and, contrary to KnowRob, it does \emph{not} provide any high-level
processing or reasoning capability (this is left to the users).

@@

These components exhibit different needs in terms of representation,
like different nominal spatial and/or temporal resolutions. For instance, a 3D
motion planner would typically use coarse 3D models of surrounding objects to
lower the computational load while planning, while a module assessing the
visibility of objects might need high-resolution models 
for accurate 3D rendering. This requirement of multiple
task-specific representations has been framed as the need for \emph{deep
representations} by Beetz~\cite{beetz2010towards}.

@@

Traditional robotic middlewares, like ROS, are not particularly well suited to
deal with these different needs: full geometric data can be represented, but is
not first-class citizen (a basic task like displaying a 3D mesh at an arbitrary
position is not particularly easy to perform with ROS, requiring the combination
of static Collada meshes, a URDF kinematic description, TF broadcasters, and a
3D visualisation tool like RViz) and critically,
representing and reasoning on alternative states of the environment is not
directly feasible.

Representing alternative states is however often highly desirable. For instance,
software components manipulating environment models typically perform better if
the models are physically consistent. However, low-level perception inaccuracies
often introduce hard-to-avoid physical inconsistencies (like detected objects
floating in the air, or wrongly inset into other objects).
Therefore, a post-process stage (for instance, using a physics simulation
engine) is needed to displace the objects seen by the robot into
physically-correct positions. Implemented with a classical approach (for
instance, using ROS TF frames), we would represent an object {\tt book} with two
frames: the original frame (\eg {\tt book\_frame\_raw}) and a second one
computed by the physics engine (\eg {\tt book\_frame\_corrected}). Such an
approach leads to the robot's 3D model being cluttered with multiple frames and
does not scale well.

@@

\subsection{Formal description}

An \uwds \emph{world} $\W^k \in \Ws$ consists in a pair (\emph{scene}, \emph{timeline}),
noted $\{\mathcal{S}^k,
\mathcal{T}^k\}$.

$\mathcal{S}^k=\{N^k, T^k\}$ is a tree representing a spatial scene (\emph{scene
graph}), with $N^k$ the set of nodes and $T^k$ the set of 6D geometric
transformations between the nodes ($|T^k| = |N^k| - 1$). $n^k_0 \in N^k$ is
called the \emph{root node} of the scene $\mathcal{S}^k$.

$\mathcal{T}^k=\{Evt^k, Sit^k\}$ is the world's \emph{timeline}, consisting in a
set of events $Evt^k$ (duration-less states) and a set of situations $Sit^k$
(states with a start time and an option end time).

\uwds \emph{operators} (noted $\op\colon \Ws^{0\ldots m} \rightarrow
\Ws^{0\ldots n}$) transform one world into another. Examples of operators
include \emph{filters} (noted $\filter$), a common type of unary, antiextensive
transformations such as $\filter\W^k \subseteq \W^k$, and \emph{generators}
(noted $\generator$), a class of nullary operators that are used to create new
worlds (for instance, from perception routines).

Finally, an \uwds \emph{network} $\mathcal{N} = \{\Ws, \Os\}$ is a directed graph (cycles are
permitted) of worlds $\Ws=\{\W^k\}, k\in0\ldots n$ and operators
$\Os=\{\op^l\}, l \in 0\ldots m$. As described in the following sections,
operators are typically implemented as independent software processes that connect at
runtime to their operands, \ie the \uwds worlds, to read and write them. These
processes are hereafter called \emph{clients} of the \uwds network.

@@

Figure~\ref{fig|scene} depicts a typical \uwds topology: a graph (that happens
to be an \emph{acyclic} graph on Figure~\ref{fig|scene}, but does not have to
be in the general case) of worlds and operators connecting the worlds together.
In the following, we however tend to adopt a software engineering perspective,
in which the \uwds network is made of \emph{clients} (which implement the
operators) connected through shared data structures called \emph{worlds}.
The formal model and this engineering-oriented model are isomorphic.

@@

Software components implementing \uwds operators and accessing worlds are
called clients. Some standard clients (like a 3D visualisation tool) are
provided with the \uwds package (see section~\ref{std_clients}). Clients are
otherwise written by the end users using the \uwds client API (see
section~\ref{api}).

Clients can both read and write onto the worlds they are connected to, and
automatically see updates broadcast by other clients connected to the same
world. To ensure data consistency, worlds can have many simultaneous readers,
but only one writer at a given time.

Three specific types of clients can be distinguished: \textbf{generator
clients} which implement generator operators, and create and update worlds
(`write-only' client, like the \emph{Sensor fusion} and \emph{ROS interface}
clients in Figure~\ref{fig|scene}); \textbf{leaf clients} which on the contrary
only read worlds, without modifying them (like the \emph{Computation of spatial
relations} client on Figure~\ref{fig|scene}); and \textbf{transformers} which
transform one representation into another (like the \emph{Semantic mapping}
client on Figure~\ref{fig|scene}). \textbf{Filters} are a specfic class of
transformer clients that copy an input world into an output world, performing some
filtering operation in-between (like the client \textit{Physics-based position
correction}).

@@

Nodes in an \uwds scene graph have four possible types: \textbf{objects} that
represent concrete physical objects (typically with one or several associated 3D
meshes); \textbf{entities} that represent abstract entities like reference
frames or groups of objects; \textbf{perspectives} that represent viewpoints
of the scene (like cameras or human gaze); and \textbf{fields} that represent
scalar or vector fields (like the visibility of an object, the working space of
robot, \etc -- note that fields are not yet implemented in the current version
of \uwds, see section~\ref{futurework}).

@@

More interesting is the time complexity of distributing changes across an \uwds
network. With $n$ the number of worlds and $m$ the number of operators in a
network $\mathcal{N}$, the worst-case (when every world is a parameter of every
operator) time complexity of creating or updating a node and propagating the
change across the network is $O(n \times m)$ (this effectively corresponds to
the \uwds server performing $n \times m$ requests to notify clients of the
update).  The space complexity is the same (as clients own a full copy of the
worlds they monitor), except for mesh data whose space and time complexities are
$O(1)$ (only the server stores the mesh data).

In the common case of one client performing a full update of a single world
(with $p$ nodes) at each time step, the complexity of propagating these changes across
the network would be $O(p \times m)$.

\section{API \& Clients}

@@

    in_world = ctx.worlds["example_in"]
    out_world = ctx.worlds["example_out"]

@@

Geometric data like meshes can be loaded into a world (or rather, into the whole
network as geometric data are treated as immutable resources shared amongst all
worlds) by simply sending an array of vertices and faces to the \uwds server.
As such, any client can load 3D data that is automatically made available to the
whole network. For example, an object recognition module recognises a new
object. It sends the mesh to the \uwds server and creates a new node which
references the mesh. At the next frame update, the 3D viewer sees a new node
referencing an unknown mesh. The viewer requests the mesh data from the
server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
the node transformation is updated, the 3D viewer can re-render the node without
having to download the mesh from the server again.

@@

@@

When integrated into a filter node, visibility computation allows easy
creation of new worlds representing the estimated perspectives of the different
agents. Listing~\ref{lst|humanperspective} provides the complete code of such an
\uwds client.

\begin{listing}[h!]
\begin{pythoncode}
import copy
import underworlds
from underworlds.tools.visibility import VisibilityMonitor

with underworlds.Context("Human PoV") as ctx:

    source = ctx.worlds["base"]
    target = ctx.worlds["human_perspective"]

    # pick up the first node named 'human'
    human = source.scene.nodebyname("human")[0]
    target.scene.nodes.append(human)

    # VisibilityMonitor computes the set of visible 
    # nodes from a given viewpoint
    visibility = VisibilityMonitor(ctx, source)

    node_mapping = {}

        for node in visibility.from_camera(human):

            newnode = node.copy()

            # track the correspondence between nodes
            # in source and target scenes
            if node in node_mapping:
                newnode.id = node_mapping[node].id
            else:
                node_mapping[node] = newnode

            # reparent the nodes to the new root
            if node.parent == source.scene.rootnode.id:
                newnode.parent = target.scene.rootnode.id

            target.scene.nodes.update(newnode)

        source.scene.waitforchanges()

\end{pythoncode}
    \caption{Example of an \uwds client that creates a world named {\tt
    human\_perspective} from a source world {\tt base} by only updating objects
    that are in the field of view of the human. As such, {\tt human\_perspective} is
    a real-time model of the human perspective on the scene.}

    \label{lst|humanperspective}
\end{listing}

\section{Performances and Application Example}
\label{application}

\subsection{Performances}
\label{performances}

The main performance metric of \uwds is the \emph{propagation time}: the time is
take of a change in a given world to cascade to all the dependant

@@

Like traditional robotic middleware, \uwds offers a form of distributed
computation based on message passing. However, it distinguishes itself from
existing middleware in significant ways. Most importantly, \uwds purposefully
does not offer any \emph{general} capability to distribute computation and data
flow amongst independent components: it focusses specifically on distributing
environment models, both spatial (geometric models) and temporal (events and
situations). In that sense, \uwds really is a \emph{distributed datastructure}
with a narrow focus on the needs of spatio-temporal representations.

@@

\subsubsection{Representation capabilities} as presented in
section~\ref{design}, the current version of \uwds allows to represent
\emph{objects}, \emph{abstract entities} like groups and \emph{perspectives}.
\emph{Fields} are also part of the \uwds design, but are not yet implemented.

\paragraph*{Representation of fields} fields are commonly used to represent
continuously-valued spatial entities. Fields might or might not be spatially
bounded. Examples include the working space of a robot arm (spatially bounded),
the field of view of a camera (spatially bounded), proxemics (potentially
unbounded). We plan to represent fields in \uwds using the memory-efficient
octomaps~\cite{hornung2013octomap}. Similarly to geometric data, octomaps
will not be directly stored with the nodes (nodes will refer to them through 
handles), but unlike geometric data, they will not be treated as immutable
datasets by the server, permitting real-time updates.

@@

\subsubsection{Manipulation of representations} \uwds makes it possible to
cheaply clone and modify multiple worlds. However, little is provided to
jointly manipulate sets of worlds. In that regard, we plan to add a mechanism to
efficiently \emph{compare} two worlds. This would permit a client
to quickly identify modifications applied to a world, and could prove especially
useful when combined with temporal events to answer questions like ``what has
changed in the scene?''. Comparisons between worlds involves the definition of
an efficient fuzzy \emph{diff} algorithm for 3D scene graphs, able to
selectively ignore non-relevant micro displacements.

\subsubsection{Implementation and Integration} finally, we plan to continue to
improve the integration of \uwds into existing software architectures. A
short-term goal is to provide excellent C++ support, with a high-level,
user-friendly C++ client library. This is critical for a broader adoption of \uwds
within the robot community. Support for other languages might follow, depending
on demands and open-source contributions.

In term of implementation, we are also investigating alternative back-ends that
could compete with gRPC to provide the synchronisation mechanisms between
clients. Due to the high-level nature of \uwds (a set of shared, tree-like data
structures), the use of the GIT version control system is one particular (if
unusual) option being studied. The scene graph would be represented as a
hierarchy of directories, nodes would be files in these directories, creating
new worlds would consist in creating a branch in the repository, and sharing
would be supported through the usual pull/push GIT mechanisms. Initial
prototyping shows that under the right conditions (in particular, an in-memory
file system), performance is not an issue. Automatic conflicts resolution, on
the other hand, might prove more challenging.

@@

We have introduced \uwds, a novel framework for shared and composable
spatio-temporal representations of a robot's world. The key contributions of our
approach are: a composite data structure for environment representation within a
robotic software architecture, made of a scene graph and a timeline; a mechanism
to efficiently and transparently share this data structure amongst a set of
clients (the software modules of the robot); a cascading architecture permitting
the explicit of representation of alternative states of the world while
maintaining a network of dependencies; a typology of clients
(\emph{root}, \emph{leaf}, \emph{filter}, \emph{transformer}) to systematically
describe such a network of representations dependencies.

>>>>

However, {\sc Spark} is a monolithic module that does not support sharing its
internal 3D model with other external components. In contrast, \uwds focuses on
offering a shared and distributed representation of the environment within the
robot's software architecture. This also distinguishes \uwds from complex
cognitive toolkits like KnowRob (as found in OpenEASE~\cite{beetz2015open}).
While these tools maintain a spatio-temporal model of the world, this model is
internal and not meant to be made widely accessible to other external processes.
\uwds focuses instead on re-usability and sharing of distributed spatio-temporal
models.  As such, \uwds can be seen as a middleware for spatio-temporal world
models and, contrary to KnowRob, it does \emph{not} provide any intrinsic high-level
processing or reasoning capability. Such reasoning skills are implemented in
loosely-coupled \emph{clients} (see Section~\ref{apiclients} hereafter).

@@

These components exhibit different needs in terms of representation,
like different nominal spatial and/or temporal resolutions. For instance, a 3D
motion planner would typically use coarse 3D models of surrounding objects to
lower the computational load while planning, while a module assessing the
visibility of objects might need high-resolution models 
for accurate 3D visibility testing. This requirement of multiple
task-specific representations has been framed as the need for \emph{deep
representations} by Beetz~\cite{beetz2010towards}.

@@

Traditional robotic middlewares, like ROS, are not particularly well suited to
deal with these different needs: full geometric data can be represented, but is
not first-class citizen: a basic task like displaying a 3D mesh at an arbitrary
position is not particularly easy to perform with ROS, requiring the combination
of static Collada meshes, a URDF kinematic description, TF broadcasters, and a
3D visualisation tool like RViz. Critically, simultaneously representing and
reasoning on alternative states of the environment is not directly feasible.

Representing alternative states is however often highly desirable. For instance,
software components manipulating environment models typically perform better if
the models are physically consistent. However, low-level perception inaccuracies
often introduce hard-to-avoid physical inconsistencies (like detected objects
floating in the air, or wrongly inset into other objects).
Therefore, a post-process stage (for instance, using a physics simulation
engine) is needed to move the objects seen by the robot into
physically-correct positions. Implemented with a classical approach (for
instance, using ROS TF frames), we would represent an object {\tt book} with two
frames: the original frame (\eg {\tt book\_frame\_raw}) and a second one
computed by the physics engine (\eg {\tt book\_frame\_corrected}). Such an
approach leads to the robot's 3D model being cluttered with multiple frames and
does not scale well.

@@

%\subsection{Formal description}
%
%An \uwds \emph{world} $\W^k \in \Ws$ consists in a pair (\emph{scene}, \emph{timeline}),
%noted $\{\mathcal{S}^k,
%\mathcal{T}^k\}$.
%
%$\mathcal{S}^k=\{N^k, T^k\}$ is a tree representing a spatial scene (\emph{scene
%graph}), with $N^k$ the set of nodes and $T^k$ the set of 6D geometric
%transformations between the nodes ($|T^k| = |N^k| - 1$). $n^k_0 \in N^k$ is
%called the \emph{root node} of the scene $\mathcal{S}^k$.
%
%$\mathcal{T}^k=\{Evt^k, Sit^k\}$ is the world's \emph{timeline}, consisting in a
%set of events $Evt^k$ (duration-less states) and a set of situations $Sit^k$
%(states with a start time and an option end time).
%
%\uwds \emph{operators} (noted $\op\colon \Ws^{0\ldots m} \rightarrow
%\Ws^{0\ldots n}$) transform one world into another. Examples of operators
%include \emph{filters} (noted $\filter$), a common type of unary, antiextensive
%transformations such as $\filter\W^k \subseteq \W^k$, and \emph{generators}
%(noted $\generator$), a class of nullary operators that are used to create new
%worlds (for instance, from perception routines).
%
%Finally, an \uwds \emph{network} $\mathcal{N} = \{\Ws, \Os\}$ is a directed graph (cycles are
%permitted) of worlds $\Ws=\{\W^k\}, k\in0\ldots n$ and operators
%$\Os=\{\op^l\}, l \in 0\ldots m$. As described in the following sections,
%operators are typically implemented as independent software processes that connect at
%runtime to their operands, \ie the \uwds worlds, to read and write them. These
%processes are hereafter called \emph{clients} of the \uwds network.

@@

Figure~\ref{fig|scene} depicts a typical \uwds topology: a graph (that happens
to be an \emph{acyclic} graph on Figure~\ref{fig|scene}, but does not have to
be in the general case) of worlds, with clients connecting the worlds to each
others.
%In the following, we however tend to adopt a software engineering perspective,
%in which the \uwds network is made of \emph{clients} (which implement the
%operators) connected through shared data structures called \emph{worlds}.
%The formal model and this engineering-oriented model are isomorphic.

@@

Software components implementing accessing \uwds worlds are
called clients. 
Clients can both read and write onto the worlds they are connected to, and
automatically see updates broadcast by other clients connected to the same
world. To ensure data consistency, worlds can have many simultaneous readers,
but only one writer at a given time.

\uwds provides several standard clients (like a 3D visualisation tool or a
physics engine simulator). Clients are however typically written by the end
users, depending on the needs of one's specific architecture.

%Three specific types of clients can be distinguished: \textbf{generator
%clients} which implement generator operators, and create and update worlds
%(`write-only' client, like the \emph{Sensor fusion} and \emph{ROS interface}
%clients in Figure~\ref{fig|scene}); \textbf{leaf clients} which on the contrary
%only read worlds, without modifying them (like the \emph{Computation of spatial
%relations} client on Figure~\ref{fig|scene}); and \textbf{transformers} which
%transform one representation into another (like the \emph{Semantic mapping}
%client on Figure~\ref{fig|scene}). \textbf{Filters} are a specfic class of
%transformer clients that copy an input world into an output world, performing some
%filtering operation in-between (like the client \textit{Physics-based position
%correction}).

@@

Nodes in an \uwds scene graph have three possible types: \textbf{objects} that
represent concrete physical objects (typically with one or several associated 3D
meshes); \textbf{entities} that represent abstract entities like reference
frames or groups of objects; \textbf{perspectives} that represent viewpoints
of the scene (like cameras or human gaze).
%; and \textbf{fields} that represent
%scalar or vector fields (like the visibility of an object, the working space of
%robot, \etc -- note that fields are not yet implemented in the current version
%of \uwds, see section~\ref{futurework}).

@@

More interesting is the time complexity of distributing changes across an \uwds
network. With $n$ the number of worlds and $m$ the number of clients in an \uwds
network, the worst-case (when every world is a parameter of every
client) time complexity of creating or updating a node and propagating the
change across the network is $O(n \times m)$ (this effectively corresponds to
the \uwds server performing $n \times m$ requests to notify clients of the
update).  The space complexity is the same (as clients own a full copy of the
worlds they monitor), except for mesh data whose space and time complexities are
$O(1)$ (only the server stores the mesh data).

In the common case of one client performing a full update of a single world
(with $p$ nodes) at each time step, the complexity of propagating these changes across
the network would be $O(p \times m)$. Figure~\ref{fig|performances} shows
measured propagation time for one change across 20 cascading worlds.

\section{API \& Clients}
\label{apiclients}

@@

    in_world = ctx.worlds["world1"]
    out_world = ctx.worlds["world2"]

@@

%Geometric data like meshes can be loaded into a world (or rather, into the whole
%network as geometric data are treated as immutable resources shared amongst all
%worlds) by simply sending an array of vertices and faces to the \uwds server.
%As such, any client can load 3D data that is automatically made available to the
%whole network. For example, an object recognition module recognises a new
%object. It sends the mesh to the \uwds server and creates a new node which
%references the mesh. At the next frame update, the 3D viewer sees a new node
%referencing an unknown mesh. The viewer requests the mesh data from the
%server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
%the node transformation is updated, the 3D viewer can re-render the node without
%having to download the mesh from the server again.

@@

\subsubsection{Physics simulation}

@@

When integrated into a filter node, visibility computation allows easy
creation of new worlds representing the estimated perspectives of the different
agents.
%Listing~\ref{lst|humanperspective} provides the complete code of such an
%\uwds client.

%
%\begin{listing}[h!]
%\begin{pythoncode}
%import copy
%import underworlds
%from underworlds.tools.visibility import VisibilityMonitor
%
%with underworlds.Context("Human PoV") as ctx:
%
%    source = ctx.worlds["base"]
%    target = ctx.worlds["human_perspective"]
%
%    # pick up the first node named 'human'
%    human = source.scene.nodebyname("human")[0]
%    target.scene.nodes.append(human)
%
%    # VisibilityMonitor computes the set of visible 
%    # nodes from a given viewpoint
%    visibility = VisibilityMonitor(ctx, source)
%
%    node_mapping = {}
%
%    while True:
%
%        for node in visibility.from_camera(human):
%
%            newnode = node.copy()
%
%            # track the correspondence between nodes
%            # in source and target scenes
%            if node in node_mapping:
%                newnode.id = node_mapping[node].id
%            else:
%                node_mapping[node] = newnode
%
%            # reparent the nodes to the new root
%            if node.parent == source.scene.rootnode.id:
%                newnode.parent = target.scene.rootnode.id
%
%            target.scene.nodes.update(newnode)
%
%        source.scene.waitforchanges()
%
%\end{pythoncode}
%    \caption{Example of an \uwds client that creates a world named {\tt
%    human\_perspective} from a source world {\tt base} by only updating objects
%    that are in the field of view of the human. As such, {\tt human\_perspective} is
%    a real-time model of the human perspective on the scene.}
%
%    \label{lst|humanperspective}
%\end{listing}
%

\section{Application Example}
\label{application}

@@

Like traditional robotic middleware, \uwds offers a form of distributed
computation based on message passing. However, it distinguishes itself from
existing middleware in significant ways. Most importantly, \uwds purposefully
does not offer any \emph{general} capability to distribute computation and data
flow amongst independent components: it focusses specifically on distributing
environment models, both spatial (geometric models) and temporal (events and
situations). In that sense, \uwds really is a \emph{distributed datastructure}
that addresses the specific needs of spatio-temporal modelling,
including the modelling of hypothetical, alternative world models, something
that traditional middlewares like ROS do not address adequately.

@@

\subsubsection{Representation capabilities} as presented in
section~\ref{design}, the current version of \uwds allows to represent
\emph{objects}, \emph{abstract entities} like groups and \emph{perspectives}.
\emph{Fields} are also part of the \uwds design, but are not yet implemented.
Fields are commonly used to represent
continuously-valued spatial entities. Fields might or might not be spatially
bounded. Examples include the working space of a robot arm (spatially bounded),
the field of view of a camera (spatially bounded), proxemics (potentially
unbounded). We plan to represent fields in \uwds using the memory-efficient
octomaps~\cite{hornung2013octomap}. Similarly to geometric data, octomaps
will not be directly stored with the nodes (nodes will refer to them through 
handles), but unlike geometric data, they will not be treated as immutable
datasets by the server, permitting real-time updates.

@@

%\subsubsection{Manipulation of representations} \uwds makes it possible to
%cheaply clone and modify multiple worlds. However, little is provided to
%jointly manipulate sets of worlds. In that regard, we plan to add a mechanism to
%efficiently \emph{compare} two worlds. This would permit a client
%to quickly identify modifications applied to a world, and could prove especially
%useful when combined with temporal events to answer questions like ``what has
%changed in the scene?''. Comparisons between worlds involves the definition of
%an efficient fuzzy \emph{diff} algorithm for 3D scene graphs, able to
%selectively ignore non-relevant micro displacements.

\subsubsection{Implementation and Integration} we plan to continue to
improve the integration of \uwds into existing software architectures. A
short-term goal is to provide excellent C++ support, with a high-level,
user-friendly C++ client library. This is critical for a broader adoption of \uwds
within the robot community. Support for other languages might follow, depending
on demands and open-source contributions.

%In term of implementation, we are also investigating alternative back-ends that
%could compete with gRPC to provide the synchronisation mechanisms between
%clients. Due to the high-level nature of \uwds (a set of shared, tree-like data
%structures), the use of the GIT version control system is one particular (if
%unusual) option being studied. The scene graph would be represented as a
%hierarchy of directories, nodes would be files in these directories, creating
%new worlds would consist in creating a branch in the repository, and sharing
%would be supported through the usual pull/push GIT mechanisms. Initial
%prototyping shows that under the right conditions (in particular, an in-memory
%file system), performance is not an issue. Automatic conflicts resolution, on
%the other hand, might prove more challenging.

@@

We have introduced \uwds, a novel framework for shared and composable
spatio-temporal representations of a robot's world. The key contributions of our
approach are: a composite data structure for environment representation within a
robotic software architecture, made of a scene graph and a timeline; a mechanism
to efficiently and transparently share this data structure amongst a set of
clients (the software modules of the robot); a cascading architecture permitting
the explicit of representation of alternative states of the world while
maintaining a network of dependencies.
%; a typology of clients
%(\emph{root}, \emph{leaf}, \emph{filter}, \emph{transformer}) to systematically
%describe such a network of representations dependencies.
